---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r, echo=TRUE}
setwd('C:/Users/mtsc7/Desktop/coursera')
activity <- read.csv("activity.csv")
summary(activity)
str(activity)
```

## What is mean total number of steps taken per day?

## (a) Histogram of total number of steps taken per day
```{r, echo=TRUE}
steps_day <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
hist(steps_day$steps, xlab = "Steps each day", main = "Total number of steps taken each day", col = "blue")
```
## (b) Mean and median of the total number of steps taken per day
```{r, echo=TRUE}
mean_steps <- mean(steps_day$steps)
mean_steps <- round(mean_steps, digits = 1)
print(mean_steps)

median_steps <- median(steps_day$steps)
median_steps <- round(median_steps, digits = 1)
print(median_steps)
```
ANSWER: The mean step is `r mean_steps` and median step is `r median_steps`.

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
