---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r, echo=TRUE}
setwd('C:/Users/Lenovo/Documents/Coursera/Reproducible Research/Week 2')
activity <- read.csv("activity.csv")
summary(activity)
str(activity)
```

## What is mean total number of steps taken per day?
## (1) Calculate the total number of steps taken per day
## (2) Histogram of total number of steps taken per day
```{r, echo=TRUE}
steps_day <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
hist(steps_day$steps, xlab = "Steps each day", main = "Total number of steps taken each day", col = "blue")
```

## (3) Mean and median of the total number of steps taken per day
```{r, echo=TRUE}
mean_steps <- mean(steps_day$steps)
mean_steps <- round(mean_steps, digits = 1)
print(mean_steps)

median_steps <- median(steps_day$steps)
median_steps <- round(median_steps, digits = 1)
print(median_steps)
```
Answer: The mean step is `r mean_steps` and median step is `r median_steps`.

## What is the average daily activity pattern?
## (1) Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo=TRUE}
mean_steps_ts <- aggregate(steps ~ interval, data = activity, FUN = mean, na.rm = TRUE)
plot(mean_steps_ts$interval, mean_steps_ts$steps, type = "l", col = "blue", xlab = "Intervals", ylab = "Total steps per interval", main = "Number of steps per interval (averaged) (Exclude NA)")
```

## (2) Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=TRUE}
max_steps <-max(mean_steps_ts$steps)
max_steps_interval <- mean_steps_ts$interval[which(mean_steps_ts$steps == max_steps)]
print(max_steps_interval)

max_steps <- round(max_steps, digits = 1)
print(max_steps)
```
Answer: The maximum steps taken is `r max_steps` and occurred at interval `r max_steps_interval`.

## Imputing missing values
## (1) Calculate and report the total number of missing values in the dataset 
## (2) Devise a strategy for filling in all of the missing values in the dataset.
## (3) Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo=TRUE}
total_na <- sum(is.na(activity))

Steps_Interval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
activity_split <- split(activity, activity$interval)

#Impute missing data for each interval
for(i in 1:length(activity_split)){
  activity_split[[i]]$steps[is.na(activity_split[[i]]$steps)] <- Steps_Interval[i]
}
activity_imputed <- do.call("rbind", activity_split)
activity_imputed <- activity_imputed[order(activity_imputed$date) ,]
```
Answer: The total number of missing value is `r total_na`.

## (4) Make a histogram of the total number of steps taken per day and calculate the mean/median total number of steps taken per day.
```{r, echo=TRUE}
steps_imputed <- tapply(activity_imputed$steps, activity_imputed$date, sum)
hist(steps_imputed, xlab = "Number of Steps", main = "Number of steps per Day (After Imputed)", col = "blue")

#Mean and median number of steps taken each day
imputed_mean_steps <- mean(steps_imputed)
imputed_mean_steps <- round(imputed_mean_steps, digits = 1)

imputed_median_steps <- median(steps_imputed)
imputed_median_steps <- round(imputed_median_steps, digits = 1)

#Compare the results in a dataframe
compare_mean_median <- data.frame(c(mean_steps, median_steps), c(imputed_mean_steps, imputed_median_steps))
colnames(compare_mean_median) <- c("Before Impute NA", "After impute NA")
rownames(compare_mean_median) <- c("mean", "median")

library(xtable)
xt <- xtable(compare_mean_median)
print(xt, type  = "html")
```
ANSWER: Imputing missing values does not change the mean value (`r imputed_mean_steps`) and insignificantly increased the median value (from `r median_steps` to `r imputed_median_steps`).

## Are there differences in activity patterns between weekdays and weekends?
## (1) Create a new factor variable in the dataset with two levels – “weekday” and “weekend”.
## (2) Make a panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends.
```{r, echo=TRUE}
activity_imputed$wktype <- ifelse(weekdays(as.Date(activity_imputed$date)) == "Saturday" | weekdays(as.Date(activity_imputed$date)) == "Sunday", "weekend", "weekday")

#Average steps per interval for weekdays and weekends
steps_weekday <- tapply(activity_imputed[activity_imputed$wktype == "weekday" ,]$steps, activity_imputed[activity_imputed$wktype == "weekday" ,]$interval, mean, na.rm = TRUE)
steps_weekend <- tapply(activity_imputed[activity_imputed$wktype == "weekend" ,]$steps, activity_imputed[activity_imputed$wktype == "weekend" ,]$interval, mean, na.rm = TRUE)

#Create a 2 panel plot for weekday and weekend
par(mfrow=c(1,2))

plot(as.numeric(names(steps_weekday)), 
     steps_weekday, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Activity in Weekdays", 
     type = "l")

plot(as.numeric(names(steps_weekend)), 
     steps_weekend, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Activity in Weekends", 
     type = "l")
```

## ANSWER: Activity (by number of steps) starts earlier during weekdays over a shorter duration (by occurrences of high peaks) while activity starts later during weekends and over a longer duration.
