---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r, echo=TRUE}
setwd('C:/Users/Lenovo/Documents/Coursera/Reproducible Research/Week 2')
activity <- read.csv("activity.csv")
summary(activity)
str(activity)
```

## What is mean total number of steps taken per day?
## (2) Histogram of total number of steps taken per day
```{r, echo=TRUE}
steps_day <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
hist(steps_day$steps, xlab = "Steps each day", main = "Total number of steps taken each day", col = "blue")
```
## (3) Mean and median of the total number of steps taken per day
```{r, echo=TRUE}
mean_steps <- mean(steps_day$steps)
mean_steps <- round(mean_steps, digits = 1)
print(mean_steps)

median_steps <- median(steps_day$steps)
median_steps <- round(median_steps, digits = 1)
print(median_steps)
```
ANSWER: The mean step is `r mean_steps` and median step is `r median_steps`.

## What is the average daily activity pattern?
## (1) Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo=TRUE}
mean_steps_ts <- aggregate(steps ~ interval, data = activity, FUN = mean, na.rm = TRUE)
plot(mean_steps_ts$interval, mean_steps_ts$steps, type = "l", col = "blue", xlab = "Intervals", ylab = "Total steps per interval", main = "Number of steps per interval (averaged) (Exclude NA)")
```

## (2) Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=TRUE}
max_steps <-max(mean_steps_ts$steps)
max_steps_interval <- mean_steps_ts$interval[which(mean_steps_ts$steps == max_steps)]
print(max_steps_interval)

max_steps <- round(max_steps, digits = 1)
print(max_steps)
```
ANSWER: The maximum steps taken is `r max_steps` and occurred at interval `r max_steps_interval`.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
